\section{Introduction}

Scientific workflows are used almost universally across scientific domains for solving complex and large-scale computing and data analysis problems. The importance of workflows is highlighted by the fact that they have underpinned some of the most significant discoveries of the past decades~\cite{badia2017workflows}. Many of these workflows have significant computational, storage, and communication demands, and thus must execute on a range of large-scale platforms, from local clusters to public clouds and upcoming exascale HPC platforms~\cite{ferreiradasilva-fgcs-2017}. Managing these executions is often a significant undertaking, requiring a sophisticated and versatile software infrastructure. 

Historically, infrastructures for workflow execution consisted of complex, integrated systems, developed in-house by workflow practitioners with strong dependencies on a range of legacy technologies---even including sets of ad-hoc scripts. Due to the increasing need to support workflows, dedicated workflow systems were developed to provide abstractions for creating, executing, and adapting workflows conveniently and efficiently while ensuring portability.  While these efforts are all worthwhile individually, there are now hundreds of independent workflow systems~\cite{workflow-systems}. These workflow systems are created and used by thousands of researchers and developers, leading to a rapidly growing corpus of workflows research publications. The resulting workflow system technology landscape is fragmented, which may present significant barriers for future workflow users due to many seemingly comparable, yet usually mutually incompatible, systems that exist. 

\input{tbl_challenges_efforts}

In the current workflow research, there are conflicting theoretical bases and abstractions for what constitutes a workflow system. It may be possible to translate between systems that use the same underlying abstractions; however, the contrary is not feasible. Specifically, typical systems have a layered model that abstractly underlie them: (i)~if the models are the same for two systems, they are compatible to some extent, and if they implement the same layers, they can be interchanged (modulo some translation effort); (ii)~if the models are the same for two systems, but they are implemented by components at different layers, they can be complementary, and may have common elements that could be shared; (iii)~if the models are distinct, workflows or system components are likely not exchangeable or interoperable. As a result, many teams still elect to build their own custom solutions rather than adopt, adapt, or build upon, existing workflow systems. This current state of the workflow systems landscape negatively impacts workflow users, developers, and researchers~\cite{deelman2018future}.

The WorkflowsRI~\cite{workflowsri} and ExaWorks~\cite{al2021exaworks} projects have partnered to bring the workflows community (researchers, developers, science and engineering users, and cyberinfrastructure experts) together to collaboratively elucidate the R\&D efforts necessary to remedy the above situation. They conducted a series of virtual events entitled ``Workflows Community Summits,'' in which the overarching goal was to (i)~develop a view of the state of the art, (ii)~identify key research challenges, (iii)~articulate a vision for potential activities, and (iv)~explore technical approaches for realizing (part of) this vision. The summits gathered over 70 participants, including lead researchers and developers from around the world, and spanning distinct workflow systems and user communities. 
The outcomes of the summits have been compiled and published in two technical reports~\cite{ferreiradasilva2021wcs, wcs2021technical}. In this paper, we summarize the discussions and findings by presenting a consolidated view of the state of the art, challenges, and  potential efforts, which we synthesize into a community roadmap. Table~\ref{tab:challenges} presents, in the form of top-level themes, a summary of those challenges and targeted community activities. Table~\ref{tab:roadmap} summarizes a proposed community roadmap with technical approaches. 

The remainder of this paper is organized as follows. Sections~\ref{sec:fair}-\ref{sec:community} provide a brief state of the art and challenges for each theme and proposed community activities. Section~\ref{sec:roadmap} discusses technical approaches for a community roadmap. Section~\ref{sec:conclusion} concludes with a summary of  discussions.

\input{tbl_roadmap}
